### 2016-12-20 Supplementary information.tex
Last changes before moving to different location

\documentclass[review, authoryear]{elsarticle}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\usepackage{pdflscape}
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
\usepackage{amssymb, amsmath}
\usepackage{topcapt}
%\usepackage[round]{natbib}
\usepackage{booktabs}
\usepackage{subscript}
\usepackage{textcomp}

%%
\usepackage[colorinlistoftodos, textwidth=4cm, shadow]{todonotes}
\DeclareRobustCommand{\Carlos}{\todo[author=Carlos, inline, color=blue!40, size=\small]}
\DeclareRobustCommand{\Jhoa}{\todo[author=Jhoa, inline, color=yellow!40, size=\small]}
%%
\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\journal{Ocean \& Coastal Management}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-harv}

\begin{document}

\begin{frontmatter}

\title{Supplementary information}
\author[MPI,CYB]{Jhoanata M. Bolivar}
\author[MPI,CYB]{Carlos A. Sierra}
\address[MPI]{Max Planck Institute for Biogeochemistry, Hans-Kn\"oll-Str. 10, 07745 Jena, Germany}
\address[CYB]{Research Center on Ecosystems and Global Change Carbono \& Bosques, Medell\'in, Colombia}

\end{frontmatter}

%\date{}							% Activate to display a given date or no date

\end{document}

% Requires the booktabs if the memoir class is not being used
\begin{table}[htbp]
   \centering \tiny
   \topcaption{Above ground biomass (AGB) data sources used to estimation of total AGB carbon stock in Colombian mangroves and development of predictive models.} % requires the topcapt package
   \begin{tabular}{p{2.0cm}p{2.0cm}p{4.0cm}p{1.0cm}p{1.0cm}p{2.0cm}p{2.5cm}} % Column formatting, @{} suppresses leading/trailing space
      \toprule 
        &&&\multicolumn{2}{c}{Coordinates}&\multicolumn{1}{c}{AGB}&\\      
      Department&Coast&Location&X&Y&(Mg ha-1)&Source\\
            \midrule 
La Guajira&Caribbean&Brazo Riito-Rancher\'{i}a river delta&-72.8931&11.5578&70.98&\citet{Lema2007}\\ 
La Guajira&Caribbean&Valle de los cangrejos-Rancher\'{i}a river delta&-72.8914&11.5588&26.78&\citet{Lema2007}\\ 
Magdalena&Caribbean&CGSM-Rinconada&-74.4938&10.9615&91.40&\citet{DelaPena2010}\\ 
Magdalena&Caribbean&CGSM-Aguas Negras&-74.6075&10.8089&16.10&\citet{DelaPena2010}\\ 
Magdalena&Caribbean&CGSM-Ca\~{n}o Grande&-74.4814&10.8619&75.80&\citet{DelaPena2010}\\ 
Magdalena&Caribbean&CGSM-Luna&-74.938&10.9071&13.80&\citet{DelaPena2010}\\ 
Magdalena&Caribbean&Chengue bay- Tayrona NNP&-74.1284&11.3178&132.10&\citet{INVEMAR2007}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a} bay-Ca\~{n}o Tij\'{io} 1&-75.8378&9.3566&147.50&\citet{Bolivar_inpreparation}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a} bay-Ca\~{n}o Tij\'{io} 2&-75.8284&9.3606&186.60&\citet{Bolivar_inpreparation}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a} bay-Ca\~{n}o Palermo&-75.8423&9.3525&129.70&\citet{Bolivar_inpreparation}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a} bay-Ca\~{n}o Grande 1&-75.8505&9.3712&153.20&\citet{Bolivar_inpreparation}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a} bay-El Claval&-75.7912&9.3874&80.20&\citet{Bolivar_inpreparation}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a} bay-Ca\~{n}o  Garzal 1&-75.8563&9.382&122.80&\citet{Bolivar_inpreparation}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a} bay-Ca\~{n}o  Garzal 2&-75.8588&9.3811&159.30&\citet{Bolivar_inpreparation}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a} bay-La Flotante-Ca\~{n}o Nisperal&-75.8029&9.3906&90.40&\citet{Bolivar_inpreparation}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a} bay-Vertel-Ca\~{n}o el Nene&-75.8397&9.3823&151.20&\citet{Bolivar_inpreparation}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a} bay-Ca\~{n}o Salado 1&-75.8721&9.4155&131.70&\citet{Bolivar_inpreparation}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a} bay-Ci\'{e}naga Galo&-75.8266&9.3673&101.80&\citet{Bolivar_inpreparation}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a} bay-Ostional&-75.8639&9.3961&89.30&\citet{Bolivar_inpreparation}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a} bay-La Zona, Rinc\'{o}n el grillo&-75.8384&9.397&72.00&\citet{Bolivar_inpreparation}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a} bay-La Camaronera&-75.7914&9.3844&74.00&\citet{Bolivar_inpreparation}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a} bay-Ci\'{e}naga Remediapobres&-75.8435&9.3679&133.20&\citet{Bolivar_inpreparation}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a}  bay-Ci\'{e}naga Soledad&-75.8464&9.3407&171.40&\citet{Bolivar_inpreparation}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a} bay-Ca\~{n}o Garzal 3&-75.8447&9.3954&102.10&\citet{Bolivar_inpreparation}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a} bay-Ca\~{n}o Grande 2&-75.854&9.3690&220.80&\citet{Bolivar_inpreparation}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a} bay-Jes\'{u}s Primera&-75.8439&9.3784&128.50&\citet{Bolivar_inpreparation}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a} bay-Ca\~{n}o Salado 2&-75.8276&9.4183&69.30&\citet{Bolivar_inpreparation}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a} bay-Angostura&-75.5885&9.4221&246.90&\citet{Bolivar_inpreparation}\\ 
C\'{o}rdoba&Caribbean&Cispat\'{a} bay-Ca\~{n}o el Soldado&-75.8548&9.3557&77.70&\citet{Bolivar_inpreparation}\\ 
Antioquia&Caribbean&Atrato river delta&-77.1005&8.0508&178.60&\citet{Blanco2012}\\ 
Antioquia&Caribbean&Puerto Cesar - Punta Coquito&-76.7407&7.9592&41.60&\citet{Blanco2012}\\ 
Antioquia&Caribbean&Punta Yarumal-Punta Las Vacas&-76.7478&8.1111&61.60&\citet{Blanco2012}\\ 
Antioquia&Caribbean&Punta Yarumal-Punta Las Vacas 2&-76.7478&8.1111&35.00&\citet{Blanco2012}\\ 
Antioquia&Caribbean&Rionegro cove 1&-76.9292&8.5458&21.20&\citet{Blanco2012}\\ 
Antioquia&Caribbean&Rionegro cove 2&-76.9292&8.5458&43.80&\citet{Blanco2012}\\ 
Antioquia&Caribbean&Rionegro cove 3&-76.9292&8.5458&30.80&\citet{Blanco2012}\\ 
Valle del Cauca&Pacific&M\'{a}laga bay-Luisico&-77.2148&4.0678&109.60&\citet{Malaga2015}\\ 
Valle del Cauca&Pacific&M\'{a}laga bay-Luisico-Winul&-77.2055&4.0842&45.30&\citet{Malaga2015}\\ 
Valle del Cauca&Pacific&M\'{a}laga bay-Luisico-Cangrejal&-77.2051&4.0874&295.90&\citet{Malaga2015}\\ 
Valle del Cauca&Pacific&M\'{a}laga bay-El Morro-Aserr\'{i}o&-77.1927&4.0506&4.00&\citet{Malaga2015}\\ 
Valle del Cauca&Pacific&M\'{a}laga bay-Corozal&-77.2678&4.0805&63.40&\citet{Malaga2015}\\ 
Valle del Cauca&Pacific&M\'{a}laga bay-Gegenera&-77.266&4.0543&51.00&\citet{Malaga2015}\\ 
Valle del Cauca&Pacific&M\'{a}laga bay-Valencia&-77.2523&4.1069&184.90&\citet{Malaga2015}\\ 
Valle del Cauca&Pacific&M\'{a}laga bay-La Estancia&-77.2714&4.1035&77.00&\citet{Malaga2015}\\ 
Valle del Cauca&Pacific&M\'{a}laga bay-Mayordomo-Manglar blanquito&-77.301&4.0420&107.50&\citet{Malaga2015}\\ 
Valle del Cauca&Pacific&M\'{a}laga bay-Caracas&-77.268&3.9871&117.20&\citet{Malaga2015}\\

               \bottomrule
   \end{tabular}
   \label{tab:SourceAGB}
\end{table}

 %Requires the booktabs if the memoir class is not being used
\begin{table}[htbp]
   \centering \tiny
  \topcaption {Statistical regression models for AGB. When log is the natural logarithm; AGB is the above ground biomass (Mg/ha); BIO9 is the mean temperature of driest quarter (\textcelsius); BIO10 mean temperature of warmest quarter (\textcelsius); BIO11 mean temperature of coldest quarter (\textcelsius); BIO16 is the precipitation of the wettest quarter (mm); EVI is the enhanced vegetation index; Lat is the absolute value of latitude (decimal degrees);  n is the number of observations; R$_a^2$  is the adjusted coefficient of determination; MSE is the mean squared error; F is the F-statistic calculated; AIC is the akaike information criterion }  % requires the topcapt package
   \begin{tabular}{p{8cm}p{0.5cm}p{1.0cm}p{1.0cm}p{1.0cm}p{1.0cm}}%Column formatting, @{} suppresses leading/trailing space
      \toprule
       Model  & n & R$_{a}^{2}$ & MSE & F & AIC\\
      \midrule
    $(1)AGB=-2.210$^{e+03}$+2.946\$^{e+00}$BIO10+4.687$^{e+00}$BIO11+2.108${e-01}$BIO16-2.334$^{e-01}$BIO17+2.465$^{e+02}$EVI$&40&0.04163 &4157&1.34&\\
    $(2)AGB=-1.876$^{e+03}$+7.766$^{e+00}$BIO1+7.64$^{e-02}$BIO4+5.413$^{e-03}$BIO12-2.908$^{e-00}$BIO15$&43&-0.0278&4338&0.72&\\
    $(3)AGB=-3.260$^{e+03}$+1.233$^{e+01}$BIO1+1.61$^{e-01}$BIO4+1.934$^{e-02$}BIO12-3.805$^{e+00}$BIO15+2.835$^{e+02}$EVI$&40&0.07695&4003&1.65&\\    
    $(4)AGB=-1.920$^{e+03}$-7.292$^{e+01}$Lat+7.878$^{e+00}$BIO10+1.846$^{e+00}$BIO11-1.115$^{e-01}$BIO16+2.404$^{e-02}$BIO17+1.561$^{e-02}$EVI$&40&0.1288&3779&1.961&\\   
    $(5)AGB=-1.372$^{e+03}$+6.358$^{e+00}$BIO10+1.789$^{e+00}$BIO11-7.389$^{e-02}$BIO16-1.229$^{e-01}$BIO17-7.746$^${e+01}Lat$&43&0.154&3571&2.529&\\   
    $(6)AGB=-1.286$^{e+03}$-9.594$^{e+01}$Lat+8.621$^{e+00}$BIO1+1.161$^{e-01}$BIO4-6.145$^{e-02}$BIO12-4.687$^{e-01}$BIO15$&43&0.1819&3453&2.868&\\    
$(7)AGB=-2.219$^{e+03}$-7.703$^{e+01}$Lat+1.113$^{e+00}$BIO1+1.885$^{e-01}$BIO4-4.071$^{e-02}$BIO12-1.735$^{e-01}$BIO15+1.813$^{e+02}$EVI$&41&0.1762&3573&2.39&\\
$(8)AGB=35338.8860-155.9330|Lat|-2490.0405\frac{BIO1}{10}+49.1282(\frac{BIO1}{10})^2-85.6399\frac{BIO11}{10}-0.1171BIO12+0.4483BIO15$&43&0.1975&3387&2.72&479.88247\\
$(9)AGB=60064.4482-127.7343|Lat|-4376.4264\frac{BIO1}{10}+81.8955(\frac{BIO1}{10})^2-17.3961\frac{BIO11}{10}-0.1014BIO12+1.9823BIO15+320.5646EVI$&41&0.2835&3046&3.26&454.33614\\
$(10)AGB=-1916.5321+7.3424BIO9+0.1126BIO17$&43&0.1456&3607&4.58&479.11018\\
$(11)AGB=-1.693$^{e+03}$-2.897$^{e+01}$Lat+7.567$^{e+00}$BIO9-4.108$^{e-02}$BIO17$&43&0.2072&3346&4.66&476.80085\\
$(12)AGB=-1.485$^{e+03}$-1.662$^{e+00}$Lat+5.443$^{e+00}$BIO9+5.658$^{e-02}$BIO17+2.770$^{e+02}$EVI$&41&0.2755&3080&4.80&452.36112\\
$(13)AGB=-1368.411-11.799Lat+5.451BIO9+231.860EVI$&41&0.2889&3023&6.42&450.71764\\
$(14)AGB=-716.047+2.522BIO9+335.402EVI$&41&0.2713&3098&8.44&450.81526\\
$(15)AGB=-1860.211+7.958BIO9-22.944Lat$&43&0.2225&3282&7.01&475.05212\\
$(16)AGB=-423.235+1.943BIO9$&43&0.01697&4149&1.72&484.20182\\
$(17)AGB=-856.79518-6.53917BIO11+0.01171BIO16+9.62033BIO9+200.26658EVI$&41&0.2859&3036&5.00&451.76863\\
$(18)AGB=-1364.4334+4.8624BIO11+0.0324BIO16+358.2106EVI$&46&0.2492&3192&5.42&452.94654\\
$(19)AGB=-822.9927+3.0227BIO11+0.2433BIO16-0.4077BIO17$&43&0.04791&4019&1.70&484.67591\\
$(20)AGB=-319.78271-13.78940BIO11-0.00399BIO16+9.78050BIO9+4.98395BIO1+231.90575EVI$&41&0.2792&3064&4.10&452.99699\\
$(21)AGB=4.173$^{e+04}$-1.606$^{e+02}$\frac{BIO11}{10}-4.733$^{e-02}$BIO16+1.602$^{e+02}$\frac{BIO9}{10}-3.083$^{e+03}$\frac{BIO1}{10}+5.685$^{e+01}$(\frac{BIO1}{10})^2+2.826$^{e+02}$EVI$&41&0.3096&2935&3.99&452.03760\\
$(22)AGB=-3.485$^{e+02}$+1.954$^{e-03}$BIO10^2+3.104$^{e-03}$BIO11^2-6.339$^{e-05}$BIO16^2+3.609$^{e-04}$BIO17^2+5.595$^{e+02}$EVI^2$&41&0.233&3261&3.43&455.54124\\
$(23)\log{AGB}=-54.802662+5.751671\log{BIO10}+5.097236\log{BIO11}-0.003928\log{BIO16}-0.004322\log{BIO17}+1.691458\log{EVI}$&41&0.2164&0.585&3.21&101.88486\\
$(24)AGB=-3301.918+718.549\log{BIO10}-122.657\log{BIO11}+30.615\log{BIO16}-6.271\log {BIO17}+131.200\log{EVI}$&41&0.2008&3397&3.01&457.22750\\
$(25)\log{AGB}=-100.7499+18.7172\log{BIO9}+0.2023\log{BIO16}+1.1821\log{EVI}$&41&0.3038&0.520&6.82&95.31651\\
$(26)\log{AGB}=-25.6202+12.1786\log{BIO9}-4.1606\log{BIO16}+0.9976\log{EVI}-1.1664|Lat|$&41&0.3777&0.465&7.07&91.59005\\
$(27)\log{AGB}=26.5733+29.8163\log{BIO9}-3.8958\log{BIO16}+0.7998\log{EVI}-0.9563|Lat|-27.7092\log{BIO11}$&41&0.3701&0.470&5.70&92.93683\\ 
$(28)AGB=1098.74-284068.51\frac{1}{BIO9}+35004.17\frac{1}{BIO16}-36.37\frac{1}{EVI}+713.59\frac{1}{|Lat|}$&41&0.256&3163&4.44&453.44744\\
$(29)\frac{1}{AGB}=1.763$^{e-01}$-6.966$^{e-04}$BIO9+3.451$^{e-05}$BIO16-1.008$^t{e-01}$EVI+5.315$^{e-03}$|Lat|$&41&0.1325&0.0014&2.53&-147.62232\\
$(30)\frac{1}{AGB}=-0.12246+28.07379\frac{1}{BIO9}-10.38792\frac{1}{BIO16}+0.01322\frac{1}{EVI}+0.18235\frac{1}{|Lat|}$&41&0.0917&0.0014&2.01&-145.73673\\
$(31)\log{AGB}=55.0753+0.3973\log{BIO9}-5.0320\log{BIO16}+2.0524\log{EVI}-8.5249\log{|Lat|}$&41&0.3756&0.466&7.01&91.73306\\
$(32)\log{AGB}=-1.615$^{e+01}$+7.970$^{e-02}$\log{BIO9}-3.738$^{e-04}$\log{BIO16}+2.816$^{e+00}$\log{EVI}-2.259$^{e-01}$|Lat|$&41&0.3193&0.508&5.69&95.26750\\
$(33)AGB=2577.5+186.7\log{BIO9}-325.0\log{BIO16}+142.5\log{EVI}-593.8\log{|Lat|}$&41&0.3223&2881&5.76&449.62118\\
$(34)AGB=-2541.15+178.36\sqrt{BIO9}-2.78\sqrt{BIO16}+259.48\sqrt{EVI}-130.84\sqrt{|Lat|}$&41&0.2699&3104&4.70&452.67672\\
$(35)\sqrt{AGB}=-7.719$^{e+01}$+3.201$^{e-01}$BIO9-7.403$^{e-05}$BIO16+1.253$^{e+01}$EVI-5.965$^{e-01}$|Lat|$&41&0.3357&7.53&6.05&205.81366\\
$(36)\log{AGB}=20.8619-4507.8468\frac{1}{BIO9}+324.2172\frac{1}{BIO16}-0.4312\frac{1}{EVI}+5.6351\frac{1}{|Lat|}$&41&0.2717&0.544&4.73&98.04207\\ 
$(37)AGB=-1368.411+5.451BIO9+231.860EVI-11.799|Lat|$&41&0.2889&3023&6.42&450.71764\\
$(38)\log{AGB}=-107.9639+20.4502\log{BIO9}+1.1227\log{EVI}-0.5814\log{|Lat|}$&41&0.313&0.513&7.07&94.76887\\
$(39)\log{AGB}=-16.91915+0.07633BIO9+3.28301EVI-0.09007|Lat|$&41&0.3323&0.499&7.63&93.60484\\
$(40)AGB=-7953.94+1482.05\log{BIO9}+82.41\log{EVI}-80.71\log{|Lat|}$&41&0.2824&3050&6.25&451.08829\\ 
$(41)AGB=-2854.91+179.82\sqrt{BIO9}+280.77|sqrt{EVI}-62.46\sqrt{|Lat|}$&41&0.2874&3029&6.38&450.80450\\
$(42)\sqrt{AGB}=-77.3401+0.3194BIO9+12.6178EVI-0.5696|Lat|$&41&0.3536&7.330&8.30&203.81456\\
$(43)\log{AGB}=25.4963-5632.1520\frac{1}{BIO23}-0.3299\frac{1}{EVI}+3.7667\frac{1}{|Lat|}$&41&0.2848&0.534&6.31&96.42112\\
$(44)\log{AGB}=20.4877-4101.9051\frac{1}{BIO9}-0.3843\frac{1}{EVI}$ &41&0.2862&0.533&9.02&95.43342\\
$(45)\log{AGB}=-74.413+14.816\log{BIO9}+1.338\log{EVI}$&41&0.3181&0.509&10.33&93.56169\\
$(46)\log{AGB}=-11.93921+0.05398BIO9+4.07342EVI$&41&0.3378&0.494&11.20&92.35835\\
$(47)\sqrt{AGB}=-45.8449+0.1781BIO9+17.6167EVI$&41&0.3389&7.50&11.25&203.83289\\   
$(48)\log{AGB}=21.9702-4494.1151\frac{1}{BIO9}-67.5181\frac{1}{BIO16}-0.3576\frac{1}{EVI}$ &41&0.2674&0.547&5.866&97.40678\\
      \bottomrule
  \end{tabular}
   \label{tab:models}
\end{table}

\section*{References}

\bibliography{Bibliography/Bibliography_Article_Mangroves}
%\bibliographystyle{apalike}

\end{document}